Analytic Occultation Light Curves in Reflected Light

### 2020-04-22 tex/ms.tex
added io europa occ

%
\section{Introduction}
\label{sec:intro}
%
Introduction here.

%
\section{Overview}
\label{sec:overview}

Our goal in this paper is to derive analytic expressions for the
flux received by a distant observer from a non-uniform Lambert sphere
illuminated by a monochromatic source that may or may not be
occulted by a (possibly different) spherical body. This applies, for example,
to the
case of exoplanet phase curves, exoplanet secondary
eclipse (occultation) light curves, or exoplanet-moon and exoplanet-exoplanet
occultations, all seen in reflected light. We explicitly assume the reflecting
body is Lambertian, i.e., it scatters light isotropically, although with an
intensity that varies across the surface according to a spatially-dependent
albedo $A$. Throughout this paper, we will take $A$ to mean the \emph{spherical}
albedo, the fraction of power incident on a body at a given wavelength
that is scattered back out to space (in all directions).
Note that the spherical albedo
is closely related to the Bond albedo: the Bond albedo is the stellar
flux-weighted integral of $A(\lambda)$ over all wavelengths $\lambda$. For
more details, see \S4.5 of \citet{Schwartz2015}.

As in \citet{Luger2019}, we compute fluxes by first expanding the surface
in terms of spherical harmonics.
While in \citet{Luger2019} we expanded the emissivity of the surface, here
we instead expand the spherical albedo $A$.
Specifically, if $\bvec{y}$ is the vector
of spherical harmonic coefficients describing the albedo anywhere on
the surface and
$\by$ is the spherical harmonics basis (Equation~\ref{eq:by}),
the albedo $A$ at a point $(x, y)$ on the sky-projected disk of the body
is given by the dot product
%
\begin{align}
    \label{eq:albedo}
    A(x, y) = \by^\top (x, y) \, \bvec{y}
    \quad.
\end{align}
%
The flux measured from this body is the surface integral over the visible
portion of the body's projected disk of the albedo $A$
times the illumination profile $\mathcal{I}$ of the surface, given by
Lambert's law as
%
\begin{align}
    \label{eq:LambertsLaw}
    \mathcal{I}(\phi) = \mathcal{I}_0 \, \text{max}\big( 0, \cos\phi \big)
    \quad,
\end{align}
%
where $\mathcal{I}_0$ is the peak illumination (at the sub-source point),
which we discuss in more detail in \S\ref{sec:adapting-starry}.

@@

\section{Extended illumination source}
\label{sec:extended}
Planets to consider: 55 Cnc e, Kelt-9b.
The angular extent of the terminator past $\nicefrac{\pi}{2}$ is
%
\begin{align}
    \tau = \arcsin\left( \frac{\Delta R}{d} \right)
\end{align}

\section{Non-Lambertian scatterers}
\label{sec:nonlambertian}
In principle, we can just change $\mathbf{I}$.

\section{Performance}
\label{sec:performance}
%
See Figure~\ref{fig:speed}.
%
\begin{figure}[h!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/speed.pdf}
        \oscaption{speed}{%
            Evaluation time (vertical axis)
            and numerical precision (point size)
            for a single flux evaluation as a function of spherical
            harmonic degree for different methods.
            In orange we show results for the emitted light
            \starry algorithm from \citet{Luger2019} (solid: no gradient,
            dashed: with gradient), and in blue we show results for
            the reflected
            light algorithm from this paper (solid: no gradient,
            dashed: with gradient). For comparison, in purple we show
            results for discrete integration on a grid (solid) and
            numerical integration using two-dimensional Gaussian
            quadrature (dashed); neither of these include gradient
            evaluations. The reflected light algorithm is around one
            order of magnitude slower and less precise than the emitted
            light algorithm, but still ${\sim}4$ orders of magnitude faster
            and ${\sim}10$ orders of magnitude more precise than numerical
            integration.
            \label{fig:speed}
        }
    \end{centering}
\end{figure}

>>>>

%
\section{Introduction}
\label{sec:intro}
%
\xxx{A few paragraphs of introduction here. \citet{Russell1906}.}

Our goal in this paper is to derive analytic expressions for the
flux received by a distant observer from a non-uniform Lambert sphere
illuminated by a monochromatic source that may or may not be
occulted by a (possibly different) spherical body. This applies, for example,
to the
case of exoplanet phase curves, exoplanet secondary
eclipse (occultation) light curves, or exoplanet-moon and exoplanet-exoplanet
occultations, all seen in reflected light. We explicitly assume the reflecting
body is Lambertian, i.e., it scatters light isotropically, although with an
intensity that varies across the surface according to a spatially-dependent
albedo $A$. Throughout this paper, we will take $A$ to mean the \emph{spherical}
albedo, the fraction of power incident on a body at a given wavelength
that is scattered back out to space (in all directions).
Note that the spherical albedo
is closely related to the Bond albedo: the Bond albedo is the stellar
flux-weighted integral of $A(\lambda)$ over all wavelengths $\lambda$
\citep[see, e.g.,][]{Seager2010}.

As in \citet{Luger2019}, we compute fluxes by first expanding the surface
in terms of spherical harmonics.
While in \citet{Luger2019} we expanded the emissivity of the surface, here
we instead expand the spherical albedo $A$.
Specifically, if $\bvec{y}$ is the vector
of spherical harmonic coefficients describing the albedo anywhere on
the surface and
$\by$ is the spherical harmonics basis (Equation~\ref{eq:by}),
the albedo $A$ at a point $(x, y)$ on the sky-projected disk of the body
is given by the dot product
%
\begin{align}
    \label{eq:albedo}
    A(x, y) = \by^\top (x, y) \, \bvec{y}
    \quad.
\end{align}
%
The flux measured from this body is the surface integral over the visible
portion of the body's projected disk of the albedo $A$
times the illumination profile $\mathcal{I}$ of the surface, given by
Lambert's law as
%
\begin{align}
    \label{eq:LambertsLaw}
    \mathcal{I}(\phi) = \mathcal{I}_0 \, \text{max}\big( 0, \cos\phi \big)
    \quad,
\end{align}
%
where $\mathcal{I}_0$ is the peak illumination (at the sub-source point).
%

The piecewise nature of the illumination function at the day/night terminator
makes the problem of computing the visible flux particularly difficult; most
studies to date have tackled the problem numerically, by discretizing the surface
and computing the relevant integrals by summing over the visible
pixels \xxx{(cite)}. Recently, \citep{Haggard2018} developed an analytic
framework for computing light curves of unocculted bodies illuminated
by a point source. In this paper, we re-derive their solution under the
\starry framework and extend it for the first time to the case where
the body is occulted by another spherical body, which may or may not be the
illumination source.

In \citet{Luger2019}, we reduced the problem of computing the flux from
an occulted body in emitted light to a series of efficient, analytical
operations involving trigonometric functions of the position and size of the
occultor and certain complete elliptic integrals. In the case of reflected
light, however, the change in the limits of integration due to the
unilluminated night side breaks many of the symmetries that simplified
the flux calculation. In particular, the limits of integration now depend
on the solution to a quartic equation specifying the points of intersection
between the occultor and the day/night terminator, and the solution to
those integrals is now a function of \emph{incomplete} elliptic integrals.
The procedure for computing the flux is therefore significantly more complex.
%
We therefore defer all calculations to the Appendix, and devote the body of
the paper to validating and demonstrating applications of our approach.

\section{Reflected light curves in \starry}
\label{sec:validation}

\subsection{Sample light curves}
\label{sec:sample}

Figure~\ref{fig:earthmoon} shows a sample application of the algorithm
developed in this paper. The figure shows two mock occultation light curves
of the cloudless Earth as the Moon is seen to transit its disk, both computed
using \starry from an $l=25$ expansion of the map of the Earth.
The top panel shows an occultation in emitted light and is a reproduction of
Figure~7 in \citet{Luger2019}. The bottom panel shows a light curve for the
same occultation geometry, but seen instead in reflected light, with the
Sun to the top left and slightly out of the page.
The black curves correspond to the analytical solution, and the orange dots
to the brute force numerical solution, obtained by discretizing the disk
on a grid of ${\sim}10^6$ points. The solutions agree to within the
numerical precision of the brute force solution (about 100 ppm in this
case).

@@

\subsection{Performance}
\label{sec:performance}

Figure~\ref{fig:speed}

\begin{figure}[h!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/speed.pdf}
        \oscaption{speed}{%
            Evaluation time (vertical axis)
            and numerical precision (point size)
            for a single flux evaluation as a function of spherical
            harmonic degree for different methods.
            In orange we show results for the emitted light
            \starry algorithm from \citet{Luger2019} (solid: no gradient,
            dashed: with gradient), and in blue we show results for
            the reflected
            light algorithm from this paper (solid: no gradient,
            dashed: with gradient). For comparison, in purple we show
            results for discrete integration on a grid (solid) and
            numerical integration using two-dimensional Gaussian
            quadrature (dashed); neither of these include gradient
            evaluations. The reflected light algorithm is around one
            order of magnitude slower and less precise than the emitted
            light algorithm, but still ${\sim}4$ orders of magnitude faster
            and ${\sim}10$ orders of magnitude more precise than numerical
            integration.
            \label{fig:speed}
        }
    \end{centering}
\end{figure}

\section{Extensions}
\label{sec:extensions}

\subsection{Extended illumination source}
\label{sec:extended}
Planets to consider: 55 Cnc e, Kelt-9b.
The angular extent of the terminator past $\nicefrac{\pi}{2}$ is
%
\begin{align}
    \tau = \arcsin\left( \frac{\Delta R}{d} \right)
\end{align}

\subsection{Non-Lambertian scatterers}
\label{sec:nonlambertian}
In principle, we can just change $\mathbf{I}$.
